---
title: "Research Assignment"
---

###Introduction

This project seeks to analyse the relationship between political beliefs and views on regulation of business. The conventional association is that more Conservative people tend to be in favour limited government regulation in the free market economy, while Liberals believe in the role of the government in curbing excess and protecting consumers. To investigate this I looked at data from the 2016 General Social Survey, specifically the political identification of people and their views on whether there should be less regulation of business. The 2016 data is espescially useful as it coincided with a Presidential election year. The focus of this election shifted from a focus on the economy to a larger emphasis on social issues, particularly immigration and foreign policy, so it is worthwhile to investigate this issue which received a smaller amount of attention.

My hypothesis was that Liberals would favour more regulation while Conservatives would favour less. The hypothesis for the second independent variable, 2016 income, was that lower incomes would also favour more regulation on business. However, the correlation was expected to be less than that of political party and view on regulation.




============================================================



###Data

The first independent variable is Party ID. The distrubution of Liberals and Conservatives seems almost symmetrical, with slightly more Liberals. Interestingly, for both Liberals and Conservatives, the most popular option was to identify as a moderate within their respective party.

This data is measured on a scale of 0 to 7 - with 0 the most Liberal and 6 the most Conservative. 7 corresponds to Other Party, or Refused to Say.



```{r}
counts <- table(gss$partyid)
barplot(counts, main="Distribution of Political Identification", xlab="Party Identification - Liberal to Conservative", col=c("blue4", "royalblue1", "skyblue1", "seashell", "indianred1", "indianred3", "red", "slategray"))


table(gss$partyid)
```




There 

The second independent variable is 2016 Income. 

```{r}
table(gss$RINCOM16)
```

The distribtion of income was quite well spread out and varied, with the majority of respondents earning less than $50,000 in the last year. The data is not a true interval-ratio variable, and it is important to note that 51 people surveyed reported incomes of over $170,000, and the data does not specify how high their actual income turned out to be. Furthermore, 85 participants declined to provide their annual incomes, which makes this variable incomplete and less 




The dependent variable is Support for Less Government Regulation of Business. 1 is strong support for this and 5 is strongly against.

```{r}
table(gss$lessreg)
```

There were a total of 1444 respondents, and the larger proportion of people supported some reduction of government regulation. Almost 50% of people surveyed believed that the regulatory regime was too much of a burden for businesses. The single largest group was in favour of regulations, closely followed by the neutral group. Interestingly, the smallest group was that strongly against a reduction in regulation. This group was a third of the size of the next closest - the group moderately against a decrease in regulation. This means that just over 17% of those surveyed would like to see increased government regulation of business. This data appeared at first glance surprising, but it is important to note that views on regulation can vary significantly, for example people frequently support regulation on banks and Wall Street firms, yet ardently oppose regulation on small businesses and so called "Mom and Pop" businesses.






###Results

Here is how the distribution of regulatory views related to Party Indentification and 2016 Income:


Party Identification and Views on Regulation


```{r}
counts <- table(gss$lessreg, gss$partyid)
barplot(counts, main="Distribution of Views on Regulation by Political Party", xlab="Party Identification - Liberal to Conservative", col=c("red", "indianred1", "seashell", "skyblue1", "blue4"), legend = rownames(counts))
```


```{r}
chisq.test(table(gss$partyid, gss$lessreg))
reg1 <- lm(partyid ~ lessreg, data = gss)
summary(reg1)
plot(partyid ~ lessreg, data = gss)
```



Income and Views on Regulation


```{r}
counts <- table(gss$lessreg, gss$RINCOM16)
barplot(counts, main="Distribution of Views on Regulation by Income", xlab="2016 Income - Low to High", col=c("red", "indianred1", "seashell", "skyblue1", "blue4"), legend = rownames(counts))
```


```{r}
chisq.test(table(gss$RINCOM16, gss$lessreg))
reg2 <- lm(RINCOM16 ~ lessreg, data = gss)
summary(reg2)
plot(RINCOM16 ~ lessreg, data = gss)
```






###Conclusions

There is a significant correlation between Party ID and Views on Regulation, however the relationship between 2016 Income and Views on Regulation is much weaker. Consequently, we can demonstrate that there is a distincy relationship between Political Identification and Views on Regulation, but we cannot prove the hypotethesis that there is a link between Income and Views on Regulation using this data.

Looking first at Political identifcation, the data fits the proposed hypothesis linking Democrats with more government oversight and consumer protection, and Republicans with support for a free market and limited government. From the Chi Square test, we can make the argument that there is a clear correlation showing that Liberals support more regulation and consumer protection, while Conservatives favour a less burdensome set of laws for businesses. This is in line with the conventional idealogies of each of the two parties in the United States, the Democrats and the Republicans.

1
    